---
title: "Time-series forecasting of HDFC stock prices"
output: 
  html_document: 
    smart: no
---

### Project details: 


#### Domain: Retail
#### Project: Time-series forecasting of HDFC stock prices
#### Data: NSE-HDFC.csv

#### Description
- Time-series is an ordered sequence of values of a variable at equally spaced time intervals.
- It can be used to obtain an understanding of the underlying forces and structure that produced the observed data. Or to fit a model and proceed to forecasting, monitoring or even feedback and feedforward control.
- We have a daily stock prices data of HDFC from 1998 to 2017.

#### Case questions
- You have to forecast the stock prices for HDFC for the next 3 months. Apart from forecasting the prices, in this time series analysis, try to answer the following fundamental questions using the same dataset:

1. What are the components of a time series?
2. How to create lags of a time series?
3. How to decompose it?
4. How to de-trend, de-seasonalize a time series?

#### Guidelines

- Use techniques like simple moving average, Holtwinters and ARIMA.
- For this time series analysis take only the closing prices (Close) for the day into account.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading required packages

```{r}

library(forecast)
library(TTR)
library(Metrics)
library(tidyr)
library(dplyr)
library(zoo)

```

### Loading the data
### stocks are provided for years 1998 to 2017 (only working days - weekends and holidays are excluded)

```{r}

stocks <- read.csv('NSE-HDFC.csv')
str(stocks)

```

### Creating time series data with input dates

```{r}

#remove duplicates
stocks <- na.omit(stocks)

#sort by date
stocks1 <- stocks
stocks1 <- stocks1 %>% mutate(Date=as.Date(Date,"%d-%m-%Y")) %>% arrange(Date)

#time series with freq of 5
stocks1_ts <- ts(stocks1$Close, frequency = 5)

#converting to regular zoo series with freq 1/5. Here 5 is used as data is corresponding to business days only
stocks1_zr <- as.zooreg(stocks1_ts)

head(stocks1_zr)
tail(stocks1_zr)

```

Can infer that:

- 1(1) to 1(5) corresponds to "business week 1"
- 945(1) to 945(2) corresponds to "business week 945"


### Plotting the time series data

```{r}

plot(stocks1_zr)

```

### Seasonal trend decompostion


```{r}

seasonaldecom <- stl(stocks1_zr,s.window = "periodic")
seasonaldecom

```

### Plotting the Seasonal trend decompostion

```{r}
plot(seasonaldecom)
monthplot(stocks1_zr)

```

Can infer that;

- There exists seasonality and trend in data
- However seasonality has very less effect compared to trend


### Initialize the results table

```{r}
results <- NULL
results <- data.frame()

```

### Simple Moving Average 

```{r}

#install.packages('smooth')
library(smooth)

SMAi <- sma(stocks1_zr, h=65,silent="graph")

#forecasting

plot(forecast(SMAi,h=65))

# Accuracy

temp <- list(sqrt(SMAi$cf))
temp <- as.data.frame(temp,col.names = 'RMSE', row.names = 'SMA')

results <- rbind(results,temp)

results['SMA',]


```


Can infer that:

- RMSE for the model is 59.85817

### Single exponential smoothing
### Holt-Winters exponential smoothing without trend and without seasonal component.

```{r}

SES_Auto <- HoltWinters(stocks1_zr,beta = FALSE,gamma=FALSE)
SES_Auto

SES_Auto_forcast <- forecast(SES_Auto,h=65)

# Accuracy

temp <- list(accuracy(SES_Auto_forcast)[,'RMSE'])
temp <- as.data.frame(temp,col.names = 'RMSE', row.names = 'SES')

results <- rbind(results,temp)

results['SES',]

# Plot the graph

plot(SES_Auto_forcast,xlab='Business week',ylab='Closing price of stocks')
lines(SES_Auto_forcast$fitted,col='green')

```


Can infer that:

- RMSE for the model is 59.86455 

### Double exponential smoothing
### Holt-Winters exponential smoothing with trend and without seasonal component.

```{r}

DES_Auto <- HoltWinters(stocks1_zr,gamma=FALSE)
DES_Auto

DES_Auto_forcast <- forecast(DES_Auto,h=65)

# Accuracy

temp <- list(accuracy(DES_Auto_forcast)[,'RMSE'])
temp <- as.data.frame(temp,col.names = 'RMSE', row.names = 'DES')

results <- rbind(results,temp)

results['DES',]

# Plot the graph

plot(DES_Auto_forcast,xlab='Business week',ylab='Closing price of stocks')
lines(DES_Auto_forcast$fitted,col='green')

```

Can infer that:

- RMSE for the model is 60.53214


### Triple exponential smoothing
### Holt-Winters exponential smoothing with trend and additive seasonal component.

```{r}

TES_Auto <- HoltWinters(stocks1_zr)
TES_Auto

TES_Auto_forcast <- forecast(TES_Auto,h=65)

# Accuracy

temp <- list(accuracy(TES_Auto_forcast)[,'RMSE'])
temp <- as.data.frame(temp,col.names = 'RMSE', row.names = 'TES_Add')

results <- rbind(results,temp)

results['TES_Add',]

# Plot the graph

plot(TES_Auto_forcast,xlab='Business week',ylab='Closing price of stocks')
lines(TES_Auto_forcast$fitted,col='green')

```

Can infer that:

- RMSE for the model is 60.90787


### ARIMA (Auto regressive Integrated Moving Average)


```{r}

ARIMA_model <- auto.arima(stocks1_zr)
summary(ARIMA_model)

ARIMA_model_forcast <- forecast(ARIMA_model,h=65)

# Accuracy

temp <- list(accuracy(ARIMA_model_forcast)[,'RMSE'])
temp <- as.data.frame(temp,col.names = 'RMSE', row.names = 'ARIMA')

results <- rbind(results,temp)

results['ARIMA',]

# Plot the graph

plot(ARIMA_model_forcast,xlab='Business week',ylab='Closing price of stocks')
lines(ARIMA_model_forcast$fitted,col='green')

```

Can infer that:

- RMSE for the model is 59.85691

### Comparing the various time series models

```{r}

results

idx <- which(results==min(results))
cat('The model selected as the best one is:',row.names(results)[idx],'\nThe RMSE for the model is:',min(results))

```


### Create file with final prediction values from the best model

```{r}

write.csv(ARIMA_model_forcast$mean,"HDFC_forecast_3months.csv")
ARIMA_model_forcast$mean

```

### Conclusion:

- These are predicted values for next 3 months which are having 65 business days
- It is forecasted that closing stock prices will decrease and strategies need to be planned accordingly.
